\documentclass{ali-presentation}

\usepackage{fontspec}
\usepackage{amsthm}
\usepackage{graphicx}

\hypersetup{
  colorlinks=true,
  allcolors=.,
  urlcolor=magenta,
}

\title{Overconfidence in Political Behavior}
\subtitle{Pietro Ortoleva and Erik Snowberg}

\date{July 28, 2025}
\author{Alistair Pattison}

\begin{document}

\maketitle

\section{Intro}

\begin{frame}
    \frametitle{Motivation}

    % TODO: motivate 
    % idological spread exists
    % some of this may caused by imperfect processing of information
    % citizent neglect correlation and update \emph{too much}
\end{frame}

\begin{frame}
    \frametitle{Basic Model}

    % TODO: overview model (no math)
    % bayesian updating fails to explain
\end{frame}

\begin{frame}
    \frametitle{Findings}

    % TODO: highlight findings 
\end{frame}

\begin{frame}
  \frametitle{Intuition}
  \begin{quote}
    ``For example, consider a
    citizen who notes the number of people in her neighborhood who are unemployed,
    and uses this information to deduce the state of the national economy. Suppose
    further that she lives in a neighborhood with high unemployment. If the citizen
    believes that the employment status of each person is relatively uncorrelated, she
    will think she has a lot of information about the state of the national economy
    (that is, she will be overconfident) and favor generous aid to the unemployed and
    loose monetary policy. If, instead, she realizes that local unemployment has a
    common cause—say, a factory shutting down—then she will understand that she
    has comparatively little information about the national economic situation, and
    believe that although the situation is bad, it is not likely to be dire, and will support
    more moderate policies.''
  \end{quote}
\end{frame}

\section{Data}

\begin{frame}
  \frametitle{Overconfidence Measure}

  % TODO: explain measure
  % ask someone a question that has a correct answer (e.g. mass of the earth),
  % then ask how confident they are in their answer.
  % this contains both waranted confidence (a.k.a. knowledge) and unwaranted over-confidence.
  % to isolate overconfidence, the authors define
  % overconfidence := confidence - E[confidence | accuracy]
  % do this four times via a survey of $n = 3,000$, then take first principal component (to reduce meas. error)
  % use econ-specific questions (e.g. unemployment)
\end{frame}

\begin{frame}
  \frametitle{Overconfidence Measure}

  % TODO: insert screenshot of table 1 from paper
\end{frame}

\begin{frame}
  \frametitle{Other Data}

  \textit{Ideology}: (1) self-reported and (2) computed from 18 issue questions.

  \textit{Voter Turnout}: voting rolls.

  \textit{Partisan Identification}: self-report (strong/moderate/lean + Dem/Rep).

  \textit{Controls}: income, education, stock/home ownership, in union, state.

  \textit{Number of Signals}: media exposure, age.
\end{frame}

\section{Theoretical Framework}

\begin{frame}
  \frametitle{Basic Model}

  % TODO: explain section I:
  % each citizen $i$ has utility
\end{frame}

\begin{frame}
  \frametitle{Similar to, but NOT Bayesian}

  % in bayesian model, posterior should converge to the truth
  % given sufficient updates

  % this should cause
  % - overconfidence -> 0
  % - population spread -> 0
  
  % in contrast, this papers model implies
  % - overconfidence -> \infty
  % - given sufficiently large $\rho$, spread -> \infty
\end{frame}

\begin{frame}
  \frametitle{Intuition}

  Consider $\rho = 1$ and $\rho_i = 0$ (experiences are perfectly correlated, but people think they're independent). Then $n \to \infty$
  \begin{itemize}
    \item
      increases confidence without increasing information (so overconfidence increases),
    \item
      pushes people towards their biased perceptions (so spread increases).
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{In the Data}

  % TODO: figure 1
\end{frame}

\section{Results}

\subsection{Overconfidence}

\begin{frame}
  \frametitle{Overconfidence and Extremeness are Correlated}
  
  % TODO: fix this
  \begin{proposition}
    Overconfidence and ideological extremeness are positively correlated.
    This is true conditional on $n$, and independent of $n$ if $\rho$ is large enough
  \end{proposition}

  % TODO: figure 2
\end{frame}

\begin{frame}
  \frametitle{In the Data}

  Measure extremeness as absolute value of ideology after residualizing:
  $$
    |\mathcal I_i - E[\mathcal I_i \ | \ \text{socioeconomic controls}]|
  $$

  % TODO: table 3
\end{frame}

\begin{frame}
  \frametitle{Effect Sizes in Context}

  % TODO: table 4
\end{frame}

\subsection{Ideology}

\begin{frame}
  \frametitle{Voter Turnout}
\end{frame}

\end{document}
